### 2025-01-15 sections/2-methods.tex
expanded section on clustering strategies

When building a tree for a real dataset, the depth would be larger than $\log n$, which would seem to increase the cost of tree-building.
However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here.
In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.

>>>>

When building a tree for a real dataset, the depth would be larger than $\log n$, which would seem to increase the cost of tree-building.
However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here, but we explore empirical results based on forcing a balanced tree in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.

### 2025-01-15 sections/4-results.tex
expanded section on clustering strategies

In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
To help with this analysis, we also added some instrumentation to our implementation of CAKES to count the number of distance computations performed during search.
These are shown in Figure~\ref{fig:results:distance-counts}.

>>>>

In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search. These strategies are the Cartesian product of balanced vs. unbalanced clustering, and the presence vs. absence of depth-first reordering as described in Section~\ref{sec:methods:clustering:depth-first-reordering}.
To help with this analysis, we also added some instrumentation to our implementation of CAKES to count the number of distance computations performed during search. We note that this instrumentation significantly slows down the wall-clock time of search, so it would not be used in a real-world application.

These four comparisons (referred to as Ball, BalancedBall, PermutedBalancedBall, and PermutedBall, where depth-first-reordering is referred to as Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset are shown in Figure~\ref{fig:results:distance-counts}. Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse; the instances (such as Depth First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which is effectively PermutedBall, no longer do. We note that on this dataset, depth-first reordering (the Permuted plotlines) seems to have little effect.
